Recovery of deleted deep sequencing data sheds more light on the early Wuhan SARS-CoV-2 epidemic

### 2021-06-17 paper/paper.tex
full draft of Discussion

More broadly, the approach taken here suggests it may be possible to learn more about the origin or early spread of SARS-CoV-2 even without a complex international investigation.
Minimally, it should be immediately possible for the NIH to determine the date and purported reason for deletion of the data set analyzed here, since the only way sequences can be deleted from the SRA is by an e-mail request to SRA staff~\citep{SRA_deletion}.
In addition, I suggest it could be worthwhile to review e-mail records to identify other SRA deletions, which are already known to include SRR11119760 and SRR11119761~\citep{USRTK2020}.
Importantly, SRA deletions do not imply any malfeasance: there are legitimate reasons for removing sequencing runs, and the SRA houses $>$13-million runs making it infeasible for their staff to validate the rationale for all requests.
However, the current study suggests that at least in one case, the trusting structures of scientific data sharing have been abused to obscure sequences relevant to the early spread of SARS-CoV-2 in Wuhan.
Therefore, a careful re-evaluation of other routine forms of scientific communication, reporting, and data sharing could shed additional light on early events in the emergence of the pandemic.

>>>>

More broadly, the approach taken here suggests it may be possible to learn more about the origin or early spread of SARS-CoV-2 even without a complex international investigation.
Minimally, it should be immediately possible for the NIH to determine the date and purported reason for deletion of the data set analyzed here, since the only way sequences can be deleted from the SRA is by an e-mail request to SRA staff~\citep{SRA_deletion}.
In addition, I suggest it could be worthwhile to review e-mail records to identify other SRA deletions, which are already known to include SRR11119760 and SRR11119761~\citep{USRTK2020}.
Importantly, SRA deletions do not imply any malfeasance: there are legitimate reasons for removing sequencing runs, and the SRA houses $>$13-million runs making it infeasible for their staff to validate the rationale for all requests.
However, the current study suggests that at least in one case, the trusting structures of science have been abused to obscure sequences relevant to the early spread of SARS-CoV-2 in Wuhan.
Therefore, a careful re-evaluation of archived forms of scientific communication, reporting, and data sharing could shed additional light on early events in the emergence of the pandemic.
